Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-08-09 Ben_Manuscripts/hdphmm/response_to_reviewers.tex

\begin{enumerate}[label={Comment \theenumi :}, leftmargin=3.9\parindent]

    %r1_comment1
    \item \textit{The manuscript by Coscia et al. analyzes diffusion and transport of solute molecules
    in a low water content lyotropic liquid crystal using a statistical model developed on the basis 
    of molecular dynamics simulations. The purpose of the statistical model is two-fold; first to 
    extrapolate computed MSDs to longer times to obtain diffusion coefficients, and second to attempt
    to gain increased physical insight on diffusion mechanisms. These issues are important in 
    highly-confined/low-water content nanoporous materials, in which diffusion occurs through complex
    mechanisms and over long timescales. The manuscript will most likely be of interest to the community
    interested in nanoporous materials for separations.}

    Author reply: We thank the reviewer for their thoughtful assessment of our work.

    %r1_comment2
    \item \textit {Unfortunately, because there is no comparison with alternative techniques, it is 
    very difficult to evaluate this new approach relative to previous work that has examined similar 
    issues. There exist numerous other methods for extrapolating MSDs to longer time, which the 
    authors have not discussed. One such method is the “trajectory-extending kinetic Monte Carlo” 
    (TEKMC) approach of Brown and coworkers see e.g. Macromolecules 2010, 43, 9210-9214, 
    Macromolecules 2015, 48, 7346-7358, and J. Phys Chem. B, 2012, 116, 95-103. The authors should 
    add a comprehensive discussion of how their approach compares to this and similar methods. The 
    paper would be significantly improved if the authors were to perform and present TEKMC results 
    for their system, so that one could better evaluate the performance of their proposed method.}

@@

    We have also modified the first sentence of the paragraph that follows in order to maintain
    good flow.

    Original text:
    \begin{quote}
    We have already advanced our understanding of LLC monomer design in our previous work
    by building stochastic time series models based on the chemical intuition gained
    from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
    membrane.
    \end{quote}

    Modified text:
    \begin{quote}
    Based on the chemical intuition gained from our previous qualitative MD studies of 
    solute transport in an H\textsubscript{II} phase LLC membrane, we have contributed our
    own stochastic modeling approaches with goals similar to both TEKMC and DBP.
    \end{quote}

    %MRS: tweaked below.
    We agree that a full comparison of our method to other methods for computing long timescale 
    diffusion such as TEKMC would be very interesting and instructive but feel it would be better 
    suited as a separate study where we could include a number of approaches, including perhaps
    recent approaches we have developed within our research group.~\cite{coscia_capturing_2020}
    Such a large-scale comparison would take more space and time than would be appropriate for 
    the current paper. Although method such as TEKMC seems a fairly straightforward approach to
    implement, it is not immediately clear how we could optimally adapt the methodology to a 
    monoclinic unit cell with two dimensional order. 
%    the grids that divide the homogeneous cubic unit cells of the referenced TEKMC studies to 
%    our system which uses a monoclinic unit cell and has 2 dimensional order.

    As stated by the reviewer, there are two goals of our work: to obtain long timescale MSDs and
    to understand mechanisms leading to diffusive behavior. TEKMC may be a suitable alternative 
    approach to obtaining our first goal, but it is not clear that it could give the same level
    of mechanistic insight that the HDP-AR-HMM offers. Although the work by Hanson et al.
    (J. Phys Chem. B, 2012, 116, 95-103) is able to provide physical insight into the effect of
    nanoparticles on diffusion of gaseous penetrants in a polystyrene polymer matrix, it does so
    by testing hypotheses motivated by observations of the diffusion constant (predicted by TEKMC)
    in response to different temperatures and nanoparticle concentrations. This is actually pretty
    in line with the type of analysis we performed in our initial LLC transport work (J. Phys Chem. B, 2019, 123, 6314-6330)
    where we used the mean squared displacement and the qualitative characteristics of the solute
    time series in order to hypothesize transport mechanisms. In this work we "re-discover" these
    mechanisms with minimal human intervention and characterize them in detail using the 
    HDP-AR-HMM while \textit{simultaneously} providing a method to estimate diffusion constants.

@@

    %r1_comment3
    %BJC: not sure if I should ignore the JCTC part completely or basically just say what the editor said.
    %MRS: my response below (Basically, saying what the editor said, in case it is send back out).
    \item \textit{Much of the writing seems more suitable for JCTC rather than JPC. I encourage the
    authors to better emphasize and discuss “new physical insight” and “chemical details”, and spend
    less time on methodology. For example pages 5-16 are all on methodology, yet it is not until 
    Figure 8 of the paper (page 26) that the authors give any description of what the LLC system 
    they are studying is actually composed of. Presumably the LLC is made up of self-assembled 
    monomers (in water) shown in Figure 8a). I would encourage the authors to describe their 
    system more concretely much earlier in the paper... there is also no information on system 
    size, box dimensions, etc.}

    Author reply: We thank the reviewer for encouraging us to include more physical insight and
    further details about our molecular simulations. We received feedback from the editor along 
    with these comments that the scope of JPC has been expanding, and that the methodological 
    details are appropriate. We have studied the structure of and transport of solutes within 
    this particular LLC system extensively in our previous work.~\cite{coscia_chemically_2019}
    We feel strongly that our current work should emphasize the HDP-AR-HMM as a more general 
    tool for understanding complex dynamics in systems beyond our example system. We would prefer
    not to cut any methodological details, as this would hurt the paper's reproducibility, and
    lengthening the paper by including additional physical insight may narrow the audience willing
    to dive in.

    %TODO    
    %BJC: probably could add a picture of the system. Or is it okay to just reference another
    % paper? I would probably use the same pic anyways or something very similar.
    %MRS: never bad to include pictures of the system, so go ahead.

@@

    We do agree that there are certain basic simulation details that we have neglected to 
    mention. Therefore we have made the following edits to section 2.1 of the methods:

    We added the highlighted text to the first sentence of the section:
    \begin{quote}
      We studied transport of solutes in the H\textsubscript{II} phase using \hl{our
      previously developed} atomistic molecular model of four pores in a 
      \hl{$9.4\times9.4\times7.6$ nm} monoclinic unit cell with 10\% water by weight, 
      \hl{totaling 62,000 atoms}.    
    \end{quote}

    We have also added the following text describing the model:
    \begin{quote}
    We modeled water molecules using the TIP3P water model,~\cite{jorgensen_comparison_1983}
    We used the Antechamber module shipped with AmberTools18 to parameterize monomers and 
    solutes using the Generalized Amber Force Field (GAFF).~\cite{wang_development_2004} We 
    assigned charges using the AM1-BCC model as implemented in Antechamber.~\cite{jakalian_fast_2000,jakalian_fast_2002}
    GAFF is very commonly used in simulations with diverse organic chemical functionality, 
    and was developed for consistency within the AMBER ecosystem, which uses TIP3P as the
    water model. We chose these standard force fields and generalizable parameterization 
    approaches so that a range of solutes and monomers could be self-consistently and easily 
    studied. The main goal of this work is to develop stochastic models which can reproduce
    solute behavior on MD time scales, independent of the choice of force field. In future 
    work, it may be beneficial to employ specialized force fields, including polarizable 
    force fields or other higher-order effects in order to improve the accuracy of structure 
    parameters and selectivity predictions.
    \end{quote}

@@

    %r1_comment4
    \item \textit{The authors should more clearly note that their discussion largely applies only 
    to very low water content LLC regimes (such as the 10\% LLC studied). For higher water content 
    LLCs (e.g. $>$ 15-20\%) it is straightforward to obtain the diffusion limit from standard MD 
    simulations. While the authors are surely well aware of this, the reader may not be, and it is
    important to clearly distinguish the regimes.}

    Author reply: We thank the reviewer for making this point. 
    %BJC2: modded based on MRS suggestion    
    In systems with less obstructions to motion, such as LLCs with a higher water content, most 
    solutes and water will be relatively unimpeded, and diffusion constants can be obtained by 
    standard techniques.    
    However, we do contend that our method would still be very useful for studying
    solute dynamics in these higher water content systems 
    %BJC2: incorporated MRS suggestion with small modifications    
    by allowing one to study the dynamics of solutes that have longer residence times and interact
    with moieties in a complex medium. %portions of the system away from the pore center.
    %Interactions characterized by different dynamical behavior even on short time scales can
    %still be highly informative. 
    This is the basis for the point we make in the introduction:``...direct calculations of diffusion 
    constants may obscure the molecular mechanisms controlling transport". Therefore we would
    prefer not to limit the range of application of this method. 
    %BJC2: added    
    However, we do agree that it would be helpful to the reader to clarify that the level of modeling 
    which we present in this work is certainly not merited in cases where they are only interested in
    diffusion constants. Therefore, we have modified a paragraph in the introduction in order to
    encourage readers to pursue simpler approaches if the system allows it.

    Updated text with additions highlighted in yellow:

    In this work, we are particularly interested in the design of lyotropic liquid 
    crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
    phases can be cross-linked into mechanically strong membranes capable of highly 
    selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
    are characterized by hexagonally packed, uniform-sized and straight pores, an ideal 
    geometry for high throughput transport. Their pores are lined with the LLC monomer's
    functional groups which can potentially be designed to interact with solutes in a 
    chemically-specific manner. It would be extremely useful to experimentalists if we 
    could connect LLC monomer design with measurable quantities such as diffusivity and
    selectivity.\hl{In many systems, including LLC membranes with water contents higher
    than the 10 wt\% water system studied here, making this connection can be relatively 
    straightforward calculation with standard techniques since small solutes and water 
    can move relatively unimpeded and therefore enter a diffusive regime on timescales 
    accessible to MD. When possible, we encourage the reader to take advantage of the 
    simplest approach possible. However, for the system that has capture our interest,} 
    this goal is complicated by the non-Brownian hopping and trapping behavior induced 
    by the membrane's diverse topology and inhomogeneous structure resulting in subdiffusive
    behavior out to multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}

    %r1_comment5
    \item \textit{In the HDP-AR-HMM method, it is not clear to me how the solute molecules are 
    constrained to be confined within the pore radius (which they must be on physical grounds 
    based on steric interactions). From Figure 2, it “looks” like this happens to work out, but
    for more complicated systems (e.g. gyroid LLCs with 3D channels) such a constraint may be 
    necessary and it is unclear how the method would generalize. This should be discussed.}

@@

    Author reply: We thank the reviewer for posing this question as this a constraint to which 
    we gave thought but neglected to emphasize in the main text. 
    %BJC: he's not saying that solutes should stay in the aqueous pore region right? We talk pretty
    % extensively about solute behavior in the tails so his steric argument is confusing me.

    %MRS: It is a little confusing. First, we never use the term "pore
    %radius", since the structure of the pore makes it very hard to
    %define a single pore radius.  The term "pore region" is used once
    %(about line 995), that that should probably be "pore center".
    %BJC2: changed pore region to pore center
    %Second, it's not clear what is meant when they say it's not clear
    %how they are constrained when they must be by steric
    %interactions.  Since steric interactions do NOT keep them in the
    %pore region (at least for the smallest solutes).  Figure 2 is a selection of a trajectory.

    %    First, we would like to clarify
%    that our simulations suggest that small solutes can partition out of the pores and into
%    the dense tail region. This is not a particularly favorable move as evidenced by our radial
%    distribution functions, but it does happen and could conceivably lead to rare jumps of solutes
%    between pores (although we have not observed any).

    %MRS: how about this?  
    %BJC2: seems good. made some minor edits
    We are grateful for the opportunity to clarify some of the details of the simulation 
    geometries. We avoid defining a ``pore radius'' because the radial distribution of 
    head groups and even tail groups is spread over several nanometers and we have
    observed clear evidence that solutes can similarly partition radially throughout 
    the pores.~\cite{coscia_understanding_2019,coscia_chemically_2019} 
%    This partitioning occurs to the extent that smaller molecules like water and methanol
%    can in fact move between pores in rare instances.
    %BJC2: I don't think I've actually published anything stating or proving the above. 
    %I think it's particularly hard to get by the crosslinks. But I do believe it could happen.
    %rephrased with that in mind.
    Therefore, it may be possible that, in rare instances, smaller molecules like water and
    methanol may move between pores. However the rarity of these events and the fact that 
    they occur orthogonal to the transport dimension means that they do not significantly 
    affect transport, and can be interpreted mathematically as simply going deep into the 
    dense tail region before returning to the same pore.

    Because of this, our stochastic simulations are constrained to remain close to a single 
    pore by the trajectory generation procedure. Each state is parameterized by a mean in 
    the radial direction computed based on the solute's distance from the \textit{nearest 
    pore center}, making it impossible for the solute to get further than approximately half
    the pore-to-pore distance away (the exact maximum distance is an an angle-dependent 
    quantity due to the hexagonally packed pore configuration).
    %MRS: though I'm not QUITE sure that's right, since the hexagonal packing means that the transition between pores
    %occurs at different distances depending on the angle. 
    %BJC: Yeah, I know what you mean. I added the bit in parentheses above to acknowledge that.
    Each time a state change occurs, the mean position of the solute in the $r$ dimension 
    jumps to the appropriate mean. However, in the axial direction, we do not parameterize 
    the mean since it should be unbounded in that dimension. The $z$ mean of each segment 
    is set based on the position immediately before the state change occurred.

    The generalization of this procedure to three dimensions will be under active investigation 
    in the near future. As mentioned, we are interested in applying this approach to bicontinuous
    cubic phases like the gyroid. Based on our current understanding, we would model solute dynamics
    in three dimensions and set the mean vector, $\mathbf{c}$ equal to $\mathbf{0}$. Of course, this
    may result in solutes exploring a three dimensional space that does not share the geometric 
    features of the studied phase. Our hypothesis is that we can add one or more additional 
    dimensions to the time series passed to the HDP-AR-HMM that includes a distance from the pore centers. 
    We could define the pore center using a spline or a surface and transitions would likely
    favor moves towards states that exist closer to the pore interiors. Alternatively, it may prove
    necessary to constrain solute positions to the means in each dimension. However, these extensions
    are beyond the scope of the current paper.
